---
title: "Netflix"
format:
  html: 
    code-tools:
      source: true
      toggle: false
      caption: "Code"
execute:
    warning: false
    message: false
---

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(stringr)

netflix_titles <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-04-20/netflix_titles.csv')

love_titles <- netflix_titles |>
  filter(str_detect(title, "Christmas")) |>
  mutate(year_added = as.numeric(str_extract(date_added, "\\d{4}"))) |>
  group_by(year_added) |>
  summarise(count = n())

ggplot(love_titles, aes(x = year_added, y = count)) +
  geom_line(color = "red") +
  geom_point() +
  scale_x_continuous(breaks = seq(2000, 2021, by = 1)) +
  labs(title = "Netflix Titles Containing 'Christmas' Released Over the Years",
       x = "Year", y = "Number of Titles")
```

The graph shows a steady increase in the number of Netflix titles containing "Christmas" from 2013 and peaking in 2020. However, then there is a massive drop to nearly zero in 2021, likely due to filming delays caused by COVID-19.

```{r, echo=FALSE}
love_2019 <- netflix_titles |>
  mutate(year_added = as.numeric(str_extract(date_added, "\\d{4}"))) |>
  filter(year_added == 2020) |>
  filter(str_detect(title, "Christmas")) |>
  count(type)

ggplot(love_2019, aes(x = type, y = n, fill = type)) +
  geom_bar(stat = "identity") +
  labs(title = "Netflix Movies vs TV Shows Containing 'Christmas' Added in 2019", 
       x = "Type", y = "Count")
```
The bar graph shows whether more movies or TV shows with "Christmas" in the title were released in 2020, the year with the most releases with "Christmas" in the title on Netflix  From the graph we can see that 25 of the releases were movies and 5 were TV shows, indicating that Netflix produces way more movies than tv shows with the word "Christmas" in the title.

```{r, echo=FALSE}
christmas_words <- netflix_titles |> 
  mutate(after_christmas = str_extract(description, "(?i)(?<=Christmas )\\w+")) |>
  filter(!is.na(after_christmas)) |>
  group_by(after_christmas) |>
  summarise(n = n()) |> 
  arrange(desc(n)) |>
  head(10)

christmas_words
```
The table presents the top 10 most common words after "Christmas" in Netflix titles. The most common word is "Eve", appearing 10 times. This is followed by "special", "to", and "with", each occurring 3 times.

Sources: https://github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-04-20/readme.md
https://www.kaggle.com/datasets/shivamb/netflix-shows 
