---
title: "Final Presentation"
author: "Frederik Spors"
subtitle: "May 13th 2025"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Project 1 - Data Visualization

::: panel-tabset

### Overview

- Two TidyTuesday datasets

- Data on the total number of flights per month (2010-2018)

- Data on the change in CO2 emission

### Flight per Month

```{r, echo = FALSE}
library(ggplot2)
library(dplyr)
library(readr)

monthly_passengers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-12-24/monthly_passengers.csv')

monthly_totals <- monthly_passengers |>
  group_by(Month) |>
  summarise(total_flights = sum(Total, na.rm = TRUE)) |>
  arrange(Month)
```

```{r}
#| fig-alt: "Line graph showing the total number of flights on average in every month from 2010-2018 in 90 countries. The number of flights peak in July, probably because of holidays."

ggplot(monthly_totals, aes(x = Month, y = total_flights)) +
  geom_line(color = "red") + 
  geom_point(alpha = 0.2) +
  labs(title = "Total Number of Flights per Month",
       x = "Month",
       y = "Total Flights (in thousands)") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:12, labels = 1:12)
```

- Data from 90 countries from 2010-2018

- Number of flights peak in the summer

<p style="font-size: 0.6em; color: #555;">
  EUROCONTROL. (2025). <i>European Air Navigation Performance Review Reports</i>. Original source: 
  <a href="https://ansperformance.eu/data/" target="_blank">https://ansperformance.eu/data/</a><br>
  Accessed via TidyTuesday GitHub repository: 
  <a href="https://github.com/rfordatascience/tidytuesday/blob/main/data/2022/2022-07-12/readme.md" target="_blank">
    GitHub Repository
  </a>
</p>

### CO2 Emission

```{r, echo = FALSE}
library(ggplot2)
library(dplyr)
library(readr)

emissions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-05-21/emissions.csv')

annual_emissions <- emissions |>
  group_by(year) |>
  summarise(total_emissions = sum(total_emissions_MtCO2e, na.rm = TRUE))
```

```{r}
#| fig-alt: "Line graph showing global total emissions in MtCO2e from 1850 to 2020, with a steep and continuous rise starting around 1950 and increasing very steeply after 2000"

ggplot(annual_emissions, aes(x = year, y = total_emissions)) +
  geom_line(color = "blue", size = 1) +
  geom_point(alpha = 0.2) +
  labs(title = "Total Emissions Over The Years",
       x = "Year",
       y = "Total Emissions (MtCO2e)") +
  theme_minimal()
```

- Global emissions have increased drastically since 1950

- Caused by rapid industrialization and increased reliance on fossil fuels

<p style="font-size: 0.6em; color: #555;">
  Climate Accountability Institute. (2023). <i>Carbon Majors Database</i>. Original source: 
  <a href="https://carbonmajors.org/" target="_blank">https://carbonmajors.org/</a><br>
  Accessed via TidyTuesday GitHub repository: 
  <a href="https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-05-21/readme.md" target="_blank">
    GitHub Repository
  </a>
</p>

:::

## Comparing Traffic Violations using SQL

:::panel-tabset
### Overview
- Grouped individuals into five age brackets to compare stop distributions in Madison (WI), Mesa (AZ) and Aurora (CO)
- Investigated most common traffic violations per city to assess enforcement patterns and demographic trends
- Applied different violation-count thresholds to highlight top violations in each city, accounting for differences in population and data size

<p style="font-size: 0.6em; color: #555;">
  Pierson, E., Simoiu, C., Overgoor, J., Corbett-Davies, S., Jenson, D., Shoemaker, A., Ramachandran, V., Goel, S., et al. (2020). 
  <i>A Large-Scale Analysis of Racial Disparities in Police Stops Across the United States</i>. 
  <i>Nature Human Behaviour</i>, 1–10.
</p>

```{r, echo=FALSE}
con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

### Age Distribution


```{sql, connection=con_traffic, output.var="age_brackets_states"}
SELECT 
  state,
  CASE 
    WHEN subject_age < 18 THEN '<18'
    WHEN subject_age BETWEEN 18 AND 30 THEN '18-30'
    WHEN subject_age BETWEEN 31 AND 45 THEN '31-45'
    WHEN subject_age BETWEEN 46 AND 65 THEN '46-60'
    ELSE '60+'
  END AS age_group,
  COUNT(*) AS stop_count
FROM (
  SELECT 'Madison' AS state, subject_age
  FROM wi_madison_2023_01_26
  WHERE subject_age IS NOT NULL
  
  UNION ALL
  
  SELECT 'Mesa' AS state, subject_age
  FROM az_mesa_2023_01_26
  WHERE subject_age IS NOT NULL
  
  UNION ALL
  
  SELECT 'Aurora' AS state, subject_age
  FROM co_aurora_2023_01_26
  WHERE subject_age IS NOT NULL) AS combined
GROUP BY state, age_group
ORDER BY state, age_group;
```

```{r, echo=FALSE}
age_brackets_states <- age_brackets_states |>
  mutate(stop_count = as.numeric(stop_count))

ggplot(age_brackets_states, aes(
  x = age_group, 
  y = stop_count, 
  fill = factor(state, levels = c("Aurora", "Mesa", "Madison")))) +
  geom_col(position = "dodge") +
  labs(
    title = "Traffic Stops by Age Groups in Madison (WI), Mesa (AZ), and Aurora (CO)",
    x = "Age Group",
    y = "Number of Stops",
    fill = "City"
  )
```



- Ages 18–30 and 31–45 had the highest number of stops in all three cities  
- Aurora had the most stops overall, especially among younger drivers  
- Madison had fewer stops across all age groups, likely due to lower traffic volume or less policing

### Frequency of Stops

**Type of Traffic Violations in Madison, WI**
```{sql, connection=con_traffic, output.var="madison_violations", echo=FALSE}
SELECT 
  violation,
  COUNT(*) AS violation_count
FROM wi_madison_2023_01_26
WHERE violation IS NOT NULL
GROUP BY violation
HAVING violation_count > 9000
ORDER BY violation_count DESC;
```
```{r, echo=FALSE}
#| fig-alt: "Table showing the most common traffic violations in Madison, WI. The top violation is 'Exceeding Speed Zones (1–10 MPH)', followed by insurance-related offenses and various speeding and signal violations."

madison_violations
```
<br>
**Type of Traffic Violations in Mesa, AZ**
```{sql, connection=con_traffic, output.var="mesa_violations", echo=FALSE}
SELECT 
  violation,
  COUNT(*) AS violation_count
FROM az_mesa_2023_01_26
WHERE violation IS NOT NULL
GROUP BY violation
HAVING violation_count > 3000
ORDER BY violation_count DESC;
```
```{r, echo=FALSE}
#| fig-alt: "Table showing the most common traffic violations in Mesa, AZ. The top violations are 'Speed Not R&P/FTC Speed to Avoid a Collision', 'Expired Registration', and 'No Proof of Insurance'."

mesa_violations
```
<br>
**Type of Traffic Violations in Aurora, CO**
```{sql, connection=con_traffic, output.var="aurora_violations", echo=FALSE}
SELECT 
  violation,
  COUNT(*) AS violation_count
FROM co_aurora_2023_01_26
WHERE violation IS NOT NULL
GROUP BY violation
HAVING violation_count > 5000
ORDER BY violation_count DESC;
```
```{r, echo=FALSE}
#| fig-alt: "Table showing the most frequent traffic violations in Aurora, CO. The most common violations include speeding, failure to present evidence of insurance, seat belt violations, and 'careless' driving."

aurora_violations
```

<br>
**Summary**

- Speeding is the most frequent violation in all three cities
- Enforcement and violation categories differ significantly
- Younger drivers (18–45) dominate traffic stops, aligning with riskier behavior trends
:::

::: {.center}

# <span style="font-size: 3em; color: #007BFF;">Thank You!</span>

<a href="https://fredsp05.github.io/FrederikSpors" target="_blank" style="
  position: absolute;
  bottom: 1.5em;
  right: 0.5em;
  background-color: #6c757d;
  color: white;
  padding: 0.5em 1em;
  border-radius: 6px;
  text-decoration: none;
  font-weight: 600;
  box-shadow: 1px 1px 4px rgba(0,0,0,0.2);
  font-size: 0.8em;
">Back to Website</a>

:::
